---
title: "Processing EvoNet output"
output: github_document
---

## Overview of important model outputs


The ouput of an EvoNet run is stored in a single data object (in all our examples it is called "evomodel" but can be any name.) This object is a list, which is a flexible R data structure where each element of a list can be any type of data object: e.g., a single number, a single character string, a vector of numbers, a matrix, a data frame, etc. We can see what objects this output list holds with the names function.

```{r chunk2, eval=F}
> names(evomodel)
[1] "param"            "control"          "nwparam"          "epi"              "stats"  
[6] "attr"             "pop"              "nw"               "coital_acts_list" "popsumm"
[11] "vl_list"          "InfMat"           "age_list"         "el"               "sessionInfo"
[16] "partner_list"     "popsumm_mats"
```

The structure of each object can be examined with the str function (see ?str). At level 0, str shows 17 objects and that the object evomodel is of class "netsim", and EpiModel class. This class attribute allows other functions to recognize the object as an EpiModel output object.

```{r chunk3, eval=F}

> str(evomodel,0)
List of 17
 - attr(*, "class")= chr "netsim"```
```

At level 1, some details of the the individual elements are shown. Here, we focus on three specific EvoNet objects: param, pop, and popsumm. The objects are coital_acts_list, partner_list, age_list, and vl_list are also EvoNet objects but discsussed in later sections.  The other objects are EpiModel specific objects and would be of interest to only very advanced users.

```{r chunk4, eval=F}

> str(evomodel,1)
List of 17
 $ param           :List of 3
 $ control         :List of 27
  ..- attr(*, "class")= chr "control.net"
 $ nwparam         :List of 1
 $ epi             :List of 3
 $ stats           : list()
 $ attr            :List of 3
 $ pop             :List of 3
 $ nw              :List of 1
 $ coital_acts_list:List of 1
 $ popsumm         :List of 3
 $ vl_list         :List of 1
 $ InfMat          :List of 1
 $ age_list        :List of 3
 $ el              :List of 3
 $ sessionInfo     :List of 3
 $ partner_list    :List of 1
 $ popsumm_mats    :List of 40
 - attr(*, "class")= chr "netsim"
```

We see that the objects param, popsumm, and pop are list of 3 elements.  In this example, 3 replicate simulations were run and for each simulation an object was stored in each of those elements. The "param" objects hold the copy of parameters used in each replicate. In 99.99% of cases, each param list element will be an exact copy of the other ones and be exactly the same as the input parameter list for the mode run (called "evoparams" in our example) To access these objects, R's list indexing is used. The input parameters are stored in the ouput object so the user can readily identify the parameterization of the model as needed.


```{r chunk5, eval=F}
#using head functio to see first several elements
head(evomodel$param[[1]])
head(evomodel$param[[2]])
head(evomodel$param[[3]])
evomodel$param[[1]]

```


The "pop" object holds the values of the individual agent attributes for the three replicates. Unlike the "param" object, each element of "pop" will be different due to each model replicate having different dynamics due to the inherent stochasticity in the model.  Each agent has has 131 separate attributes pertaining to demographic, epidemiological, treatment, and behavioral/sexual attributes.  Below is a list of these attributes. Typicall, only a small subset of these are of interest to the user.

```{r chunk6, eval=F}
> sort(names(evomodel$pop[[1]]))
  [1] "adherence_start"                "adherence_type"                 "Adherence1"                    
  [4] "Adherence2"                     "Adherence3"                     "Adherence4"                    
  [7] "age"                            "age_infection"                  "ai_prob"                       
 [10] "aim3_mutations_long"            "aim3_no_muts"                   "Aim3RoundingErrors"            
 [13] "arrival_time"                   "att1"                           "CD4"                           
 [16] "cd4_at_test"                    "CD4_at_trtmnt"                  "CD4_initial_value"             
 [19] "CD4_nadir"                      "CD4_time"                       "CD4_time_death"                
 [22] "CD4_treatment_delay_index"      "CD4count"                       "CD4tot"                        
 [25] "ChronPhase"                     "circum"                         "condom_user"                   
 [28] "CYP_6_slow"                     "d_acute"                        "diag_resist_status"            
 [31] "diag_resist_time"               "diag_status"                    "diag_time"                     
 [34] "disclosure_status"              "Donors_age"                     "Donors_age"                    
 [37] "Donors_CD4"                     "Donors_d_acute"                 "Donors_diag_status"            
 [40] "Donors_EnvirContribToLogSP0"    "Donors_Generation"              "Donors_Index"                  
 [43] "Donors_LogSetPoint"             "Donors_SetPoint"                "Donors_Total_Time_Inf_At_Trans"
 [46] "Donors_treated"                 "Donors_treated_2nd_line"        "Donors_V"                      
 [49] "Donors_ViralContribToLogSP0"    "Drug1"                          "Drug2"                         
 [52] "Drug3"                          "Drug4"                          "eligible_2nd_line_ART"         
 [55] "eligible_ART"                   "eligible_care"                  "eligible_for_prep"             
 [58] "eligible_vl_test"               "enhanced_testing"               "EnvirContribToLogSP0"          
 [61] "ever_enhanced_testing"          "Generation"                     "have_diag_partner"             
 [64] "have_disc_partner"              "have_suppressed_partner"        "I_vec"                         
 [67] "id"                             "Imm_Trig"                       "insert_quotient"               
 [70] "K"                              "known_pos_partner_duration"     "L_vec"                         
 [73] "last_disc_sex"                  "last_neg_resist_test"           "last_neg_test"                 
 [76] "LogSetPoint"                    "M_vec"                          "min_time_tx"                   
 [79] "no_partners_now_prep"           "no_partners_past_prep"          "num_consec_VL_gt1k"            
 [82] "NumRecipients"                  "on_prep"                        "OnDrug"                        
 [85] "partner_recent_test"            "pos_partner_duration"           "PPP"                           
 [88] "prep_decrease"                  "prep_list"                      "r0"                            
 [91] "rand_prob_test"                 "rand_prob_test_init"            "RandomTimeToAIDS"              
 [94] "rate_phase2"                    "role"                           "s"                             
 [97] "SetPoint"                       "sex"                            "spvl_cat"                      
[100] "sqrt_age"                       "start_aids_cd4"                 "start_max_aids"                
[103] "Status"                         "sti_status"                     "Time_Death"                    
[106] "time_hiv_sex"                   "time_hiv_sex_act"               "Time_Inf"                      
[109] "Time_Inf_Adj"                   "time_init_2nd_line"             "total_acts"                    
[112] "treated"                        "treated_2nd_line"               "tx_dropout"                    
[115] "tx_init_time"                   "tx_schedule"                    "tx_stop_time"                  
[118] "V"                              "V_vec"                          "vacc_init_time"                
[121] "vaccinated"                     "ViralContribToLogSP0"           "virus_3_plus_drug_muts"        
[124] "virus_3_plus_drug_muts"         "virus_part_res_drug"            "virus_sens_drug"               
[127] "virus_sens_vacc"                "vl_at_test"                     "vl_expected"                   
[130] "vl_peak_agent"                  "vl_phase2_trans" 
```

Each element of pop is as long as the total number of agents that existed during the model run (for a particular replicate.)  If a model starts with 300 agents and 117 enter the model by the end of the simulation, then each element of pop will consist of vectors of length 317. In the example below, the number of total agents in each replicate differed.

```{r chunk7, eval=F}
> length(evomodel$pop[[1]]$Status)
[1] 417
> length(evomodel$pop[[2]]$Status)
[1] 416
> length(evomodel$pop[[3]]$Status)
[1] 412
```

As an aside, using the lapply function is often useful for interacting with R lists (see ?lapply).
```{r chunk8, eval=F}
> unlist(lapply(1:3,function(x) length(evomodel$pop[[x]]$Status)))
[1] 417 416 412
```


The attribute "Status" refers to the current condition of an agent: 0, alive and uninfected; 1, alive and infected; -1.5, aged out of model (removed from model as maximum age was reached); -1, died of non-AIDS cause; -2, died of AIDS. The first 6 values for each Status vector can be viewed using the head function and the differences between replicates are clear.

```{r chunk9, eval=F}
> lapply(1:3,function(x) head(evomodel$pop[[x]]$Status))
[[1]]
[1] -1.5 -1.5 -2.0 -1.5  0.0  0.0

[[2]]
[1] -1.5 -1.5 -2.0 -1.5  1.0 -2.0

[[3]]
[1] -1.5 -1.5  0.0 -1.5  1.0 -2.0
```

To view all the values for replicate 1
```{r chunk10, eval=F}
evomodel$pop[[1]]$Status
```

Regular R functions can be applied to these vectors
```{r chunk11, eval=F}
> out=evomodel$pop[[1]]$Status
> table(out)
out
  -2 -1.5   -1    0    1 
  51   96    9  239   22 
```

The pop structure is such that each position in an attribute vector corresponds to the same agent. 
```{r chunk12, eval=F}
> out=evomodel$pop[[1]]
> index=10
> out$sex[index]
[1] "m"
> out$age[index]
[1] 55.0016
> out$Status[index]
[1] -1.5
> out$Time_Inf[index]
[1] NA
> out$V[index]
[1] NA
> out$LogSetPoint[index]
[1] NA
```

Agent 80 is an example of infected agent. Note that time of infection "Time_Inf" is in days since model start and V (viral load) is raw value and SPVL ("LogSetPoint") is on log10 scale.

```{r chunk13, eval=F}
> out=evomodel$pop[[1]]
> index=80
> out$sex[index]
[1] "m"
> out$age[index]
[1] 40.46227
> out$Status[index]
[1] 1
> out$Time_Inf[index]
[1] 4097
> out$V[index]
[1] 38006.65
> out$LogSetPoint[index]
[1] 4.0613
```


This shows that agent 10 was a male, that aged out (reached default maximum age at 55 years and Status of -1.5) and was never infected (Time_Inf = NA) and therefore did not have a viral load (V), SPVL (LogSetPoint) value (both were NA).



# Summary statistics

At the end of each daily timestep, a suite of summary statistics is caclulated to quantify epidemioligical, behavioral, and network dynamics of the population.  These statistics are based on individual agent attributes (e.g., infection status to calculate prevalence) and provide a time series of statistics that can be used to describe and quanitfy the evolution of the epidemic. 